Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-14 paper/agree.tex

The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components. The AGREE specifications generated by the transforms for the high-assurance components are explained. The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component. This meaning is what must be preserved by the synthesis.

@@

The AGREE specification for the SW component in the example with the added cyber requirements is given in \figref{fig:sw}. Without going into too much detail, the language is a first order predicate calculus. The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic. Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the top-level model being verified. The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component and strong enough to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output). These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself). These are the unexpanded results at the bottom of the figures.

@@

The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}. The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.

The value of \texttt{previous} in the current time step relies on values from the previous time step. The \texttt{->} operator is for initialization as the previous time step is undefined in the first step of the system. The left operand to the operator is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system. The right operand is the value of \texttt{previous} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

@@

As noted previously, the original system fails to guarantee the cyber requirement. BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor. The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation. The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}. The policy of the high assurance component can be stated directly in the wizard, or leave it  blank. In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}. 
Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.  Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run anytime throughout the design phase to determine whether those artifacts are valid.
%Additionally, the filter can be attached to any requirement the vulnerability analysis. This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021]. The \emph{OK} button creates the component and inserts it accordingly into the implementation

The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a). The guarantee is stylized for synthesis and completely defines the meaning of the output under all possible input. 
%A similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed. 
The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised and is one of the several options available in the dialogue. The definition for \texttt{policy} is taken by the system designer from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

At this point it is important to state the meaning of a high-assurance component more formally. Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component such as what might be sent over a wire by the communication fabric. An \emph{environment}, $\theta: \lval \mapsto \konst{string}$, divides the raw string into its constituent parts as given by its data definition in AADL. As such, it binds \emph{L-values} to strings where an L-value is a concept from compilers and represents anything that can appear on the left-hand side of an assignment. In the context of this presentation, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc. The environment maps any L-value in the scope of the AGREE contract specification to its associated raw representation as a subset of bytes in a contiguous string.

@@

The set $\Theta_s$ is the universe of environment bindings for the given specification $s$. 
The meaning of an AGREE contract specification is now defined as follows.
%
\[
  \mathcal{L}(s) = \{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots\}  
\]
%
The meaning is the prefix closed set of finite environment streams on which the specification is invariant. Each environment in the stream is total meaning that it binds all L-values in the scope if the specification.

This paper claims that a synthesized high-assurance component generates $\mathcal{L}(s)$ precisely. Let $\konst{synth}(s)$ be the CakeML synthesized component, and let $\theta^\prime = \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that evaluates the synthesized component on the input stream to produce a corresponding output stream that is total in its bindings. Let $I_s(\theta\ldots)$ be a function that maps an environment stream to a new environment stream that only retains the input bindings of each environment in the stream. This paper claims the following.
%
\[
  \forall \theta\ldots \in \mathcal{L}(s)\ (\konst{eval}\ \konst{synth}(s)\ I_s(\theta\ldots) = \theta\ldots) \\
\]
%

>>>>

The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components. The AGREE specifications generated by the transforms for the high-assurance components are then explained. The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component. This meaning is what must be preserved by the synthesis.

@@

The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}. The AGREE specification language is a first-order predicate calculus. The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic. Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the top-level model being verified. The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input as well as the guarantees on the system output). These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself). These are the unexpanded results at the bottom of the figures.

@@

The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since}, meaning that either the policy holds or the alert is sounding. The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}. The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.

The value of \texttt{previous} in the current time step relies on values from the previous time step. The \texttt{->} operator designates initialization, as the previous time step is undefined in the first step of the system. The left operand to the operator is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system. The right operand is the value of \texttt{previous} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

@@

As noted previously, the original system fails to guarantee the cyber requirements. BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor. The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation. The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}. The policy of the high-assurance component can be stated directly in the wizard, or it can be left  blank. In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}. 
Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.  Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run at any time to determine whether those artifacts are valid.
%Additionally, the filter can be attached to any requirement the vulnerability analysis. This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021]. The \emph{OK} button creates the component and inserts it accordingly into the implementation

The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a). The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input. 
%A similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed. 
The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.  This behavior is one of the several options available in the dialogue. The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}. As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

At this point it is important to state the meaning of a high-assurance component more formally. Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component, such as what might be sent over a wire by the communication fabric. An \emph{environment}, $\theta: \lval \mapsto \konst{string}$, divides the raw string into its constituent parts as given by its data definition in AADL. As such, it binds \emph{L-values} to strings where an L-value (a concept from compilers) represents anything that can appear on the left-hand side of an assignment. In the context of this paper, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc. The environment maps any L-value in the scope of the AGREE contract specification to its associated raw representation as a subset of bytes in a contiguous string.

@@

The set $\Theta_s$ is the universe of environment bindings for the given specification $s$. 
The meaning of an AGREE contract specification is now defined as follows.
%
\[
  \mathcal{L}(s) = \{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots\}  
\]
%
The meaning is the prefix closed set of finite environment streams on which the specification is invariant. Each environment in the stream is total, meaning that it binds all L-values in the scope if the specification.

We claim that a synthesized high-assurance component generates $\mathcal{L}(s)$ precisely. Let $\konst{synth}(s)$ be the CakeML synthesized component, and let $\theta^\prime = \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that evaluates the synthesized component on the input stream to produce a corresponding output stream that is total in its bindings. Let $I_s(\theta\ldots)$ be a function that maps an environment stream to a new environment stream that only retains the input bindings of each environment in the stream. We then claim the following:
%
\[
  \forall \theta\ldots \in \mathcal{L}(s)\ (\konst{eval}\ \konst{synth}(s)\ I_s(\theta\ldots) = \theta\ldots) \\
\]
%
